---
title: "Literature survey/ Literature Review/ Literature Search"
subtitle: ""  
author: 
  - "Dr Thiyanga S. Talagala"
  - ""
date: ''
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#081d58",
  secondary_color = "#ce1256",
  inverse_header_color = "#FFFFFF",
  base_font_size = "20px",
  text_font_size = "2rem"
)
```

# What is the purpose of a literature review?

- To learn what has already been done

- To identify gaps and opportunities

- To ensure you're not just repeating what others have already done

- To familiarize yourself with the current state of knowledge on your topic

---

# When writing your thesis/ paper

you should explain your reader how your work related to existing research and what new insights it will contribute.

---

# How can I access published research papers?

-  Academic databases and search engines

- Speak with the librarian at your institution to find out if you have access to any subscription databases

---

# Free databases/ search engines

- [Google scholar](https://scholar.google.co.uk/)

- [Directory of Open Access Journal (DOAJ)](https://doaj.org/)

- [PubMed](https://pubmed.ncbi.nlm.nih.gov/)

-  [Bielefeld Academic Search Engine (BASE)](https://www.base-search.net/)

- [E-Theses Online Service (EThOS)](https://ethos.bl.uk/Home.do;jsessionid=C76F4980B7E2A7CC3504DF016BB66100)


---

# Free databases/ search engines


- [Social Science Research Network (SSRN)](https://www.ssrn.com/index.cfm/en/)

-  [WorldWideScience](https://worldwidescience.org/)

- [Semantic Scholar](https://www.semanticscholar.org/) - “intelligent” academic search engine. Machine learning algorithm is used to prioritize the most important research

- [Public Library of Science (PLOS)](https://plos.org/)

<!--https://proofed.com/writing-tips/10-free-research-journal-databases/-->

---

# Free databases/ search engines



- [Microsoft Academic Search](https://www.semanticscholar.org/topic/Microsoft-Academic-Search/432350)


- [Dimensions](https://www.dimensions.ai/resources/type/publications/)

---

## Databases that are available through subscription

- [Scopus](https://www.scopus.com/home.uri)

- [Web of science](https://mjl.clarivate.com/search-results)

For more information: https://guides.loc.gov/library-science/subscription-databases

---

If you can access the full text of a paper, it could be because of a number of reasons, for example

- Open access

- Authenticated user

- Author-archived copy

---

# Dealing with paywalls

- Speak with the librarian

- Write to the corresponding author

- Use unpaywall to see if there is an open access version of the full text of the paper

---

# What is unpaywall?

Browser extension that will automatically point out full text or open-access version of research papers


Example: https://unpaywall.org/

---

# Preprints and repositories

- arXiv

- bioRxiv

- Institutional repositories

- OpenDOAR website

---

## Research access initiatives

- [Research4Life](https://www.research4life.org/about/programs/)

Research4Life provides low- and middle-income countries with access to scientific and professional resources (for example research that is published in a wide range of journals, etc.) through five programs: Hinari, AGORA, OARE, ARDI and GOALI

- [INASP](https://www.inasp.info/)

---

### Reading a research paper

1. Is the content clear?

2. What is the research question or hypothesis?

3. Whether the methodology is clear, understandable, replicable and relevant to achieve the research question?

4. Do methods corresponds with data analysis?

5. Answer to the research question and hypothesis?

---

## Reading papers to identify opportunities


1. What is my area of research?

2. What is known so far?

3. What is not known so far in this area?

4. Which unknowns do I need to focus in my research?

---

## Identify relevant unkowns

1. Why it is important to address a particular unkown or gap?

2. What contribution will it make to the field?

3. Do you have knowledge, expertise, equipments, time and budget to address the unkown?

After that you can formulate your research question. It should be clear and specific.

---

## Question: Your turn

1. What types of pesticides are honey bees likely to be exposed to in major crops?

2. Are honey bees exposed to pesticides in major crops?

3. What happened when bees are exposed to pesticides?

---

## Planning your literature review

Visit here: https://writeonline.ca/litreview.php?content=section1
