---
title: "Research Ethics"
subtitle: ""  
author: 
  - "Dr Thiyanga S. Talagala"
  - ""
date: ''
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#081d58",
  secondary_color = "#ce1256",
  inverse_header_color = "#FFFFFF",
  base_font_size = "20px",
  text_font_size = "2rem"
)
```


## Ethical violations in research/ Research misconduct

- Intentional plagiarism

- Potential conflict of interest

- Unintentional plagiarism

- Data falsification

- Data fabrication

- Duplicate publication

---

## Ways of handling fraud

- Rejected

- Retracted

Example:https://retractionwatch.com/retracted-coronavirus-covid-19-papers/

---

## Obtain research approval

- Ethical Review Board

---

## Sharing data

- Include all the data that is **relevant** to your research question

- Data sharing: supplementary data, share data in a public domain

- Check data sharing guidelines

---

## Ethics in writing

- Give credit to others work

- Disclose the limits on your information

- Use credible sources (Peer-reviewed journals and academic textbooks are considered to be credible sources.)


For more information: https://www.kent.edu/stark/ethical-writing-reliable-sources#:~:text=Ethical%20writing%20is%20writing%20with,racist%2C%20homophobic%2C%20etc.)

---

## Salami Slicing

- Breaking up a large study into two or more publications

---

## Submitting to journals

- Do not submit the paper to multiple journals at the same time.

---
## Other

- Authorship order (Before asking for assistance, tell your colleague about the authorship or acknowledgement)

- Supervisory authorship

- Referencing

- Conflict of interest

